---
title: 'MA 331: Homework 01'
author: "Stephaan Silne"
date: "9/2/2020"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1.

(i). Histograms and Pie Charts of X and Y


  * Histogram of X

```{r hw01_data, echo = FALSE}
bin <- seq(0,10,1)
hist(x = c(0.2,1.2,0.9,2.2,3.2,0.3,1.7,3.1,2.3,1.5,2.5,3,2.6,9), breaks = bin, xlab = "X", main = "Histogram of X", xlim = c(0,10),ylim = c(0,5) )
```

  * Histogram of Y


```{r, echo = FALSE}
hist(x = c(1.1,2.3,1.1,3.6,0.1,1.0,6.9,4.8,6.5,7.8,8.0,9.4,9.8), breaks = bin, xlab = "Y", main = "Histogram of Y", xlim = c(0,10), ylim = c(0,5) )
```

  * Pie Chart of X
    
    
```{r, echo = FALSE}
slices <- c(3,3,4,3,0,0,0,0,0,1)
lbls <- c("0-1","1-2","2-3","3-4","9-10")
pie(slices,labels = lbls, main = "Pie Chart of X")
```


  * Pie Chart of Y


```{r, echo = FALSE}

slices <- c(1,3,1,1,1,1,2,1,1,2)
lbls <- c("0-1","1-2","2-3","3-4","4-5","5-6","6-7","7-8","8-9","9-10")
pie(slices,labels = lbls, main = "Pie Chart of Y")
```


|           The distribution of $X$ is positively or right skewed, with most of the data being jumbled towards the left. The mean, median and mode are not equal for this set of data. In addition, there is an outlier between 8 to 9. The distribution of $Y$ is very uniform, as there does not seem to be any peaks. Due to this, the data for $Y$ is not skewed in any direction and is mostly symmetric.









---

(ii).

  * Boxplot of X

```{r, echo = FALSE}

boxplot(x = c(0.2,1.2,0.9,2.2,3.2,0.3,1.7,3.1,2.3,1.5,2.5,3,2.6,9), xlab = "X", main = "Boxplot of X")

```

  * Boxplot of Y
  
```{r, echo = FALSE}

boxplot(x = c(1.1,2.3,1.1,3.6,0.1,1.0,6.9,4.8,6.5,7.8,8.0,9.4,9.8), xlab = "Y", main = "Boxplot of Y")

```

  * Five-Number Summaries

```{r,echo = FALSE}

library(readxl)
hw01_data <- read_excel("C:/Users/Stephaan/Downloads/hw01 data.xlsx")
summary(hw01_data)
```

  
  
  * Variances

```{r, echo = FALSE}
var(hw01_data[1])
var(hw01_data[2])
```

  * Outliers



|       *LOW OUTLIERS FOR X*
|       $X < Q_{1} - 1.5 \cdot IQR$
|       $X < 1.275 - 1.5 \cdot (2.9 - 1.275)$
|       $X < -1.1625$
|
|
|       *LOW OUTLIERS FOR Y*
|       $Y < Q_{1} - 1.5 \cdot IQR$
|       $Y < 1.4 - 1.5 \cdot (7.575 - 1.4)$
|       $Y < -7.8625$
|
|       *HIGH OUTLIERS FOR X*
|       $X > Q_{3} + 1.5 \cdot IQR$
|       $X > 2.9 + 1.5 \cdot (2.9 - 1.275)$
|       $X > 5.3376$
|
|       *HIGH OUTLIERS FOR Y*
|       $Y > Q_{3} + 1.5 \cdot IQR$
|       $Y > 7.575 + 1.5 \cdot (7.575 - 1.4)$
|       $Y > 16.8375$



```{r, echo = FALSE}

outliers <- matrix(c("None", "None", "9.0", "None"), ncol = 2, byrow = TRUE )
colnames(outliers) <- c("X", "Y")
rownames(outliers) <- c("Low Outliers", "High Outliers")
outliers <- as.table(outliers)
outliers
```



---

(iii).

  * Scatter plot of $(X,Y)$

  
```{r, echo = FALSE}

x <- hw01_data$X
y <- hw01_data$Y

plot(x,y, main = "Scatter Plot of (X,Y)", xlab = "X", ylab = "Y", xlim = c(0,10), ylim = c(0,10), frame = FALSE)

#abline(lm(x~y), col="black")

```

  * Correlation Coefficient $r(x,y)$.
  


|                 $r(x,y) = \frac{1}{n-1} \sum_{i=1}^{n}(\frac{x_i -\bar{x}}{s_{x}})(\frac{y_i -\bar{y}}{s_{y}})$
  
```{r, echo = FALSE}

cor(x,y)

```

|           The linear association between X and Y is fairly moderate, given that $r(x,y) = 0.5679153$ and $r(x,y) \in [-1,1]$, it is halfway between a weak and strong correlation. The linear association is positive. 

---

(iv.) There are some outliers in the scatter plot of $(X,Y)$, which are $(3.2,0.1)$ and $(9.0,9.8)$.


```{r, echo = FALSE}

xout <- x > 3.1
plot(x[!xout],y[!xout], xlim = c(0,10), ylim = c(0,10), frame = FALSE, xlab = "X", ylab = "Y", main = "New Scatter Plot of (X,Y)") 

```

|
|
|
|
|
|       New correlation coefficient $r(x,y)$:




```{r, echo = FALSE}

cor(x[!xout],y[!xout])

```

---

(v).


   * The numerical results of the correlation coefficient from (iv) compared to (iii) is larger in value, showing that the linear association has became stronger given the removal of outliers in the dataset. 


---

(vi).
    
  * Normal QQ plots for observations of $X$ and $Y$.
  


```{r, echo = FALSE}

x <- hw01_data$X
Y <- hw01_data$Y
qqnorm(x, xlab = "X", main = "QQ Plot of X")
qqnorm(y, xlab = "Y", main = "QQ Plot of Y")


```


|       Of the two observations, the one more likely to be normally distributed would have to be $Y$, as the qq plot for $X$ is more right skewed. 





## Problem 2.



  * For $x_{1}, \cdots, x_{n}$, it can be shown that,
  
  
$\begin{aligned}
\sum_{i=1}^{n}(x_{i} - \bar{x})^2 &= \sum_{i=1}^{n}(x_{i}^2 - 2x_{i}\bar{x} + \bar{x}^2) \\
                                  &= \sum_{i=1}^{n}x_{i}^2 - 2\bar{x}\sum_{i=1}^{n}x_{i} +\sum_{i=1}^{n}\bar{x}^2 \\
                                  &= \sum_{i=1}^{n}x_{i}^2 - 2(n\frac{\sum_{i=1}^{n}}{n^2})\sum_{i=1}^{n}x_{i} +\sum_{i=1}^{n}\bar{x}^2\\
                                  &= \sum_{i=1}^{n}x_{i}^2 -2n\bar{x}^2 + n\bar{x}^2 \\
                                  &= \sum_{i=1}^{n}x_{i}^2 - n\bar{x}^2 \\
\end{aligned}$


|         Therefore,


$\begin{aligned}
\frac{1}{n}\sum_{i=1}^{n}(x_{i} - \bar{x})^2  &= \frac{1}{n}(\sum_{i=1}^{n}x_{i}^2 - n\bar{x}^2) \\
                                              &= \frac{\sum_{i=1}^{n}x_{i}^2}{n} - \frac{n\bar{x}^2}{n} \\
                                              &= \frac{1}{n} \sum_{i=1}^{n}x_{i}^2 - \bar{x}^2 \\
                                              \end{aligned}$









