---
title: "MA 331 HW 02"
author: "Stephaan Silne"
date: "9/10/2020"
output:
  html_document: default
  word_document: default
  pdf_document: default
---


## Problem 1.

i.
    
  -for $n = 20$
    
```{r, echo = FALSE}
## P(N <= 8.25)
pbinom(q = 8.25, size = 20, prob = 0.4, lower.tail = TRUE)
```

  -for $n = 30$

```{r, echo = FALSE}
## P(N <= 8.25)
pbinom(q = 8.25, size= 30, prob = 0.4, lower.tail = TRUE )
```
    
  -for $n = 50$
    
```{r, echo = FALSE}
## P(N <= 8.25)
pbinom(q = 8.25, size = 50, prob = 0.4, lower.tail = TRUE)
```
    
  -for $n = 75$
    
```{r, echo = FALSE}
## P(N <= 8.25)
pbinom(q = 8.25, size = 75, prob = 0.4, lower.tail = TRUE)
```
    
  -for $n = 100$
    
```{r, echo = FALSE}
## P(N <= 8.25)
pbinom(q = 8.25, size = 100, prob = 0.4, lower.tail = TRUE)
```


ii.

$\begin{aligned}
P(N \leq 8.25) = P(\frac{\frac{N}{n} - p}{\sqrt{\frac{p(1-p)}{n}}} \leq \frac{\frac{8.25}{20} - 0.4}{\sqrt{\frac{0.4(1-0.4)}{20}}}) &\approx \Phi(\frac{8.25 - 20(0.4)}{\sqrt{(20)(0.4)(1-0.4)}})\\ 
&\approx \Phi(0.114108866147)\\
&\approx
\end{aligned}$

```{r, echo = FALSE}
##P(N <= 8.25)
pnorm(0.114108866147, mean = 0, sd = 1, lower.tail = TRUE)
```

$\begin{aligned}
P(N \leq 8.25) = P(\frac{\frac{N}{n} - p}{\sqrt{\frac{p(1-p)}{n}}} \leq \frac{\frac{8.25}{30} - 0.4}{\sqrt{\frac{0.4(1-0.4)}{30}}}) &\approx \Phi(\frac{8.25 - 30(0.4)}{\sqrt{(30)(0.4)(1-0.4)}})\\ 
&\approx \Phi(-1.39754248594)\\
&\approx
\end{aligned}$

```{r, echo = FALSE}
##P(N <= 8.25)
pnorm(-1.39754248594, mean = 0, sd = 1, lower.tail = TRUE)
```

$\begin{aligned}
P(N \leq 8.25) = P(\frac{\frac{N}{n} - p}{\sqrt{\frac{p(1-p)}{n}}} \leq \frac{\frac{8.25}{50} - 0.4}{\sqrt{\frac{0.4(1-0.4)}{50}}}) &\approx \Phi(\frac{8.25 - 50(0.4)}{\sqrt{(50)(0.4)(1-0.4)}})\\ 
&\approx \Phi(-3.39193283149)\\
&\approx
\end{aligned}$

```{r, echo = FALSE}
##P(N <= 8.25)
pnorm(-3.39193283149, mean = 0, sd = 1, lower.tail = TRUE)
```

$\begin{aligned}
P(N \leq 8.25) = P(\frac{\frac{N}{n} - p}{\sqrt{\frac{p(1-p)}{n}}} \leq \frac{\frac{8.25}{75} - 0.4}{\sqrt{\frac{0.4(1-0.4)}{75}}}) &\approx \Phi(\frac{8.25 - 75(0.4)}{\sqrt{(75)(0.4)(1-0.4)}})\\ 
&\approx \Phi(-5.1265241636)\\
&\approx
\end{aligned}$

```{r, echo = FALSE}
##P(N <= 8.25)
pnorm(-5.1265241636, mean = 0, sd = 1, lower.tail = TRUE)
```

$\begin{aligned}
P(N \leq 8.25) = P(\frac{\frac{N}{n} - p}{\sqrt{\frac{p(1-p)}{n}}} \leq \frac{\frac{8.25}{100} - 0.4}{\sqrt{\frac{0.4(1-0.4)}{100}}}) &\approx \Phi(\frac{8.25 - 100(0.4)}{\sqrt{(100)(0.4)(1-0.4)}})\\ 
&\approx \Phi(-6.48094161111)\\
&\approx
\end{aligned}$

```{r, echo = FALSE}
##P(N <= 8.25)
pnorm(-6.48094161111, mean = 0, sd = 1, lower.tail = TRUE)
```


iii.


$\begin{aligned}
\text{abs. diff for $n = 20$}\\
|0.5955987 - 0.5454243| &= 0.050174\\
\text{abs. diff for $n = 30$}\\
|0.09401122 - 0.08112525| &= 0.01288597\\
\text{abs. diff for $n = 50$}\\
|0.0002305229 - 0.0003470073| &= 0.0001164844\\
\text{abs. diff for $n = 75$}\\
|(1.826106e^{-8}) - (1.475701e^{-7})| &= 1.2930904e^{-7}\\
\text{abs. diff for $n = 100$}\\
|(5.431127e^{-13}) - (4.557597e^{-11})| &= 4.50328573e^{-11}\\
\end{aligned}$


```{r, echo = FALSE}

plot(x = c(20,30,50,75,100), y = c(0.050174, 0.012886, 0.000116, 1.293090*10^-7, 4.503285*10^-11), xlab = "n", ylab = "errors", main = "Scatter Plot of Errors of Approximation")

```

iv. 
|             The greater the sample size, the less error there is in approximating, given there is more information on the population included in the sample $n$. More precision and less randomness occurs with larger sample sizes, thus allowing better approximation and estimation of the mean. 





## Problem 2.


- for $n = 20$


```{r, echo = FALSE}

zs <- c()
for (i in 1:100){
  smpl <- rnorm(n = 20, mean = 2, sd = 3)
  z <- (mean(smpl)-2)/(3/sqrt(20))
  zs <- c(zs,z)}

plot(density(zs), xlab = "x", ylab = "density", main = "Density of Standard Normal")
```

```{r, echo = FALSE}
chi <- c()
for (i in 1:100){
  smple <- rnorm(n = 20, mean = 2, sd = 3)
  chitst <- ((20 - 1) * var(smple))/(3**2)
  chi <- c(chi, chitst)}

plot(density(chi), xlab = "x", ylab = "density", main = "Density of Chi Sq. Statistic")

```


```{r, echo = FALSE}

plot(x = zs, y = chi, xlab = "standard normal", ylab = "chi-sq statistic", main = "Scatter Plot")
```

- for $n = 30$

```{r, echo = FALSE}

z30 <- c()
for (i in 1:100){
  smpl30 <- rnorm(n = 30, mean = 2, sd = 3)
  
  zs30 <- (mean(smpl30)-2)/(3/sqrt(30))
  
  z30 <- c(z30,zs30)}

plot(density(z30), xlab = "x", ylab = "density", main = "Density of Standard Normal")

```



```{r, echo = FALSE}

chi30 <- c()
for (i in 1:100){
  smpl30 <- rnorm(n = 30, mean = 2, sd = 3)
  chitst30 <- ((30 - 1) * var(smpl30))/(3**2)
  chi30 <- c(chi30, chitst30)}
  
plot(density(chi30), xlab = "x", ylab = "density", main = "Density of Chi Sq. Statistic")

```


```{r, echo = FALSE}

plot(x = z30, y = chi30, xlab = "standard normal", ylab = "chi-sq statistic", main = "Scatter Plot")

```



- for $n = 50$

```{r,echo = FALSE}


z50 <- c()
for (i in 1:100){
  smpl50 <- rnorm(n = 50, mean = 2, sd = 3)
  
  zs50 <- (mean(smpl50)-2)/(3/sqrt(50))
  
  z50 <- c(z50,zs50)}

plot(density(z50), xlab = "x", ylab = "density", main = "Density of Standard Normal")

```


```{r, echo = FALSE}

chi50 <- c()
for (i in 1:100){
  smpl50 <- rnorm(n = 50, mean = 2, sd = 3)
  
  chitst50 <- ((50 - 1) * var(smpl50))/(3**2)
  
  chi50 <- c(chi50, chitst50)}

plot(density(chi50), xlab = "x", ylab = "density", main = "Density of Chi Sq. Statistic")

```


```{r, echo = FALSE}

plot(x = z50, y = chi50, xlab = "standard normal", ylab = "chi-sq statistic", main = "Scatter Plot")


```




- for $n = 75$

```{r, echo = FALSE}


z75 <- c()
for (i in 1:100){
  smpl75 <- rnorm(n = 75, mean = 2, sd = 3)
  
  zs75 <- (mean(smpl75)-2)/(3/sqrt(75))
  
  z75 <- c(z75,zs75)}

plot(density(z75), xlab = "x", ylab = "density", main = "Density of Standard Normal")

```


```{r, echo = FALSE}

chi75 <- c()
for (i in 1:100){
  smpl75 <- rnorm(n = 75, mean = 2, sd = 3)
  
  chitst75 <- ((75 - 1) * var(smpl75))/(3**2)
  
  chi75 <- c(chi75, chitst75)}

plot(density(chi75), xlab = "x", ylab = "density", main = "Density of Chi Sq. Statistic")


```


```{r, echo = FALSE}

plot(x = z75, y = chi75, xlab = "standard normal", ylab = "chi-sq statistic", main = "Scatter Plot")
```



v. 

  - Based on the density plots of $\frac{\bar{X}-2}{\sqrt{3^2/n}}$ and $\frac{(n-1)S^2}{3^2}$, the probability distributions of $\frac{\bar{X}-2}{\sqrt{3^2/n}}$ becomes more normally distributed as the sample size increases. The shape of the distribution starts to flatten out towards the tails as $n$ gets larger, suggesting normalization. As for the distribution of $\frac{(n-1)S^2}{3^2}$, as the sample size increases, the distributions centers itself more. For $n = 20$, the distribution is skewed more towards the left, but for $n = 75$ the data is much for centralized and centered about the mean. 
  

vi. 
  - Based on the scatter plots of $(\frac{\bar{X}-2}{\sqrt{3^2/n}},\frac{(n-1)S^2}{3^2})$, the statistical association between $\frac{\bar{X}-2}{\sqrt{3^2/n}}$ and $\frac{(n-1)S^2}{3^2}$ is non-existent, with there being no obsrved pattern or linear association amongst the two distributions. 



## Problem 3.

For $N \sim B(n,p)$, 

$\begin{aligned}
\mu = E(N) &= \sum_{x = 0}^{n}x \frac{n!}{x!(n-x)!} \cdot p^{x} \cdot (1-p)^{n-x}\\
&= \sum_{x=1}^{n} \frac{n!}{(x-1)!(n-x)!} \cdot p^{x} \cdot (1-p)^{n-x}\\
&= np \sum_{x=1}^{n} \frac{(n-1)!}{(x-1)!(n-x)!} \cdot p^{x-1} \cdot (1-p)^{n-x}\\
\text{Let $n-x = (n-1) - (x-1)$},\\
 E[N] &= np \sum_{x=1}^{n} \frac{(n-1)!}{(x-1)!((n-1)-(x-1))!} \cdot p^{x-1} \cdot (1-p)^{(n-1)-(x-1)}\\
\text{Let $u = n-1$; so $v = x-1$},\\
E[N] &= np \sum_{x=1}^{n} \frac{(u)!}{(v)!((u)-(v))!} \cdot p^{v} \cdot (1-p)^{(u)-(v)}\\
\text{Since $\sum_{x=1}^{n} \frac{(u)!}{(v)!((u)-(v))!} \cdot a^{v} \cdot b^{u-v} = (a+b)^m$ for $a = p$ and $b = 1-p$},\\
E[N] &= np(p + (1-p))^m= np(1)^m\\
E[N] &= np\\
\end{aligned}$


## Problem 4.

For $T \sim \tau_{n}$ and by definition of $E[T]$,

$\begin{aligned}
\qquad \mathrm {E}[T] &= \int_{-\infty}^{\infty}tf(t)dt\\
\qquad \mathrm{E}[T]&=\frac{1}{\sqrt{n}}\frac{\Gamma(\frac{n+1}{2})}{\Gamma(\frac{1}{2})\Gamma(\frac{n}{2})}\int_{-\infty}^{\infty}t(1+\frac{t^2}{n})^{-\frac{n+1}{2}}dt\\
 &= 0\\
\end{aligned}$

With $E[T]$ undefined if $0 \leq n \leq 1$, for $n > 1$, the function $f(t)=t(1+\frac{t^2}{n})^{-\frac{n+1}{2}}$ is an odd function. Therefore, through the integrand, from $(-\infty, \infty)$ the probability density function approaches and is symmetrical around 0. Thus, the expected value or mean, $E[T] = 0$.





